---
title: "Climate Change"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Climate change is an ongoing issue that scientists frequently debate about. Though there are many factors which contribute to rising average temperatures across the United States, one claim is that crude oil consumption directly affects the rising temperatures. As statisticians, we understand that both variables are observational, meaning that there can be correlation between the two variables, but causation cannot be concluded. Thus, for this study, we are interested in discovering whether there is significant correlation between crude oil consumption levels and average temperature readings. The dataset 'oil-temp.csv' has been provided and contains the following variables:

year: Measurement year
crude_oil_consumption1: Total annual US crude oil imports measured in thousands of barrels
temp_change: Average change in US temperatures by degrees Fahrenheit

Given this data, our hypotheses are as follows:
 * H0:cor=0
 * Ha:cor≠0

Calculate the actual correlation coefficient between crude_oil_consumption and temp_change (5 points)
Perform a permutation test with 10,000 samples and store the results in a vector (15 points)
Create a histogram of your permutation test results and include a red line for the actual correlation value (10 points)
Calculate the p value for the observed correlation coefficient and its 95% confidence interval (10 points)
What do you conclude? (5 points)

***

Load packages.
```{r}
library(tidyverse)
library(infer)
```

***

Import data.
```{r}
oil <- read_csv('oil-temp.csv')
```

Viewing data.
```{r}
head(oil)
```

Let's calculate the true correlation between temperature and crude oil consumption.
```{r}
(samp_corr <- oil %>%
  summarize(Correlation = cor(crude_oil_consumption, temp_change)))
samp_corr <- samp_corr %>%
  pull()
```

Since the correlation is between oil consumption and temperature on a yearly basis, if we
randomize temperatures, it will break the correlation and create data that fits our null
hypothesis. I'll do this using the 'tidyverse' by creating 10,000 samples of our original
data, then permuting the 'temp_change' variable for each sample then calculating the
correlation between 'temp_change' and 'crude_oil_consumption' for each sample.
```{r}
perm_corrs <- oil %>%
  rep_sample_n(size = nrow(oil), reps = 10000) %>%
  group_by(replicate) %>%
  mutate(perm_temp = sample(temp_change)) %>%
  summarize(Correlation = cor(crude_oil_consumption, perm_temp))
```

Plot the correlation values under the null hypothesis and include a red verticle line at
the correlation we calculated from our actual sample.
```{r}
perm_corrs %>%
  ggplot(aes(x = Correlation)) +
  geom_density() +
  geom_vline(xintercept = samp_corr, col = 'red') +
  labs(y = 'Density',
       title = 'Distribution of Correlation Values Under Null Hypothesis')
```

Calculate the p-value and 95% confidence interval for the observed correlation coefficient.
```{r}
perm_corrs %>%
  summarize(P_Value = mean(Correlation >= samp_corr),
            Lower_Bound = P_Value - qnorm(0.975) * sqrt(P_Value * (1 - P_Value) / n()),
            Upper_Bound = P_Value + qnorm(0.975) * sqrt(P_Value * (1 - P_Value) / n()))
```

Since our null hypothesis is that the correlation is not equal to zero, we are conducting a two sided test. To be significant, we would need to have a p-value of less 0.025 (alpha = 0.05) Based on this permutation test, we can conclude that the observed correlation between
crude oil consumption and average yearly temperature change is greater than 0 at a 
statistically significant level since the p-value is much smaller than 0.025 and the 
95% confidence interval that we calculated does not contain the cutoff value of 0.025.