---
title: "Seatbelts"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The built-in R dataset Seatbelts contains details about driver deaths and injuries in Great Britain from Jan 1969 to Dec 1984. More information about the data can be accessed by running ?Seatbelts. Use the following line of code to load this data into R as a data.frame:

This data contains 8 columns and 192 rows of data. The law column is either a 0 or 1 indicating whether or not a law was in place requiring a seatbelt to be worn. We want to know if this law had an impact on the proportion of drivers who were killed out of those who were seriously injured. This proportion is calculated by dividing DriversKilled by drivers.

* Calculate the observed difference in proportion of drivers killed when the law was in
effect and when it was not in effect (10 points)
* Calculate 10,000 bootstrapped differences in proportions and store the results in a vector (15 points)
* Based on the results from step 2, calculate a 95% confidence interval on the difference in proportion of drivers killed (10 points)
* Based on your results, did the law improve the proportion of injured drivers who were killed? (5 points)

***

Load packages.
```{r}
library(tidyverse)
library(infer)
```

***

Load data.
```{r}
sblts <- as_data_frame(Seatbelts)
```

View the data.
```{r}
head(sblts)
```

I'll calculate the observed difference in proportion of drivers killed when the law was in
effect and when it was not. In this case, the 'diff()' function is subtracting the proportion of drivers killed when the law was not in effect from the proportion of drivers that were killed when the law was in effect.
```{r}
(actual_diff <- sblts %>%
  group_by(law) %>%
  summarize(Proportion_Killed = sum(DriversKilled) / sum(drivers)) %>%
  summarize(Diff_Prop_Killed = diff(Proportion_Killed)))

actual_diff <- actual_diff %>%
  pull()
```

Calculating 10,000 bootstrapped differences in proportions. The code groups by law (0 or 1)
and then takes a random sample with replacement of all the rows in each group. Then it
calculates the difference in proportion of drivers killed before and after the law change.
```{r}
boot_diffs <- sblts %>%
  select(DriversKilled, drivers, law) %>%
  rep_sample_n(size = nrow(.), reps = 10000) %>%
  group_by(replicate, law) %>%
  sample_frac(size = 1, replace = TRUE) %>%    
  summarize(Proportion_Killed = sum(DriversKilled) / sum(drivers)) %>%
  summarize(Diff_Prop_Killed = diff(Proportion_Killed))
```

Calculate a 95% confidence interval on the difference in proportion of drivers killed.
```{r}
(ci <- boot_diffs %>%
  summarize(Lower_Bound = quantile(Diff_Prop_Killed, probs = 0.025),
            Upper_Bound = quantile(Diff_Prop_Killed, probs = 0.975)))
```

Visualize the confidence interval.
```{r}
boot_diffs %>%
  ggplot(aes(x = Diff_Prop_Killed)) +
  geom_density() +
  geom_vline(xintercept = c(ci$Lower_Bound, ci$Upper_Bound), col = 'red') +
  labs(x = 'Difference',
       y = 'Density',
       title = 'Distribution of Differences in Proportion of Drivers Killed')
```

Based on the confidence interval from the bootstrap samples, we cannot conclude that the
law improved the proportion of drivers that are killed in crashes since the 95% confidence
interval around the difference in proportion of drivers killed contains 0.
